---
title: "MLB Player Scouting Report"
format:
  html:
    theme: cosmo
    toc: false
    code-fold: false
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
params:
  player_name: "Aaron Judge"
  player_type: "Hitter"
  season: 2025
  project_dir: "."
---

```{r setup, include=FALSE}
library(tidyverse)
library(gt)
library(scales)

project_dir <- normalizePath(params$project_dir)

hitters <- readRDS(file.path(project_dir, "data", "processed", "hitters.rds"))
pitchers <- readRDS(file.path(project_dir, "data", "processed", "pitchers.rds"))

player_name_param <- params$player_name
player_type <- params$player_type

get_percentile <- function(data, metric, player_value, higher_is_better = TRUE) {
  vals <- data[[metric]]
  vals <- vals[!is.na(vals)]

  pct <- mean(vals <= player_value) * 100

  if (!higher_is_better) {
    pct <- 100 - pct
  }

  round(pct, 1)
}

make_hitter_scouting <- function(data, selected_player) {
  player <- data |> filter(player_name == selected_player) |> slice(1)

  tibble(
    Metric = c("WAR", "wRC+", "OPS", "OBP", "SLG"),
    Value = c(player$war, player$w_rc_plus, player$ops, player$obp, player$slg),
    Percentile = c(
      get_percentile(data, "war", player$war),
      get_percentile(data, "w_rc_plus", player$w_rc_plus),
      get_percentile(data, "ops", player$ops),
      get_percentile(data, "obp", player$obp),
      get_percentile(data, "slg", player$slg)
    )
  )
}

make_pitcher_scouting <- function(data, selected_player) {
  player <- data |> filter(player_name == selected_player) |> slice(1)

  tibble(
    Metric = c("WAR", "ERA", "FIP", "K/9", "BB/9"),
    Value = c(player$war, player$era, player$fip, player$k_9, player$bb_9),
    Percentile = c(
      get_percentile(data, "war", player$war),
      get_percentile(data, "era", player$era, higher_is_better = FALSE),
      get_percentile(data, "fip", player$fip, higher_is_better = FALSE),
      get_percentile(data, "k_9", player$k_9),
      get_percentile(data, "bb_9", player$bb_9, higher_is_better = FALSE)
    )
  )
}

if (player_type == "Hitter") {
  player <- hitters |> filter(player_name == player_name_param) |> slice(1)
  scouting_df <- make_hitter_scouting(hitters, player_name_param)
} else {
  player <- pitchers |> filter(player_name == player_name_param) |> slice(1)
  scouting_df <- make_pitcher_scouting(pitchers, player_name_param)
}

top_metric <- scouting_df |> arrange(desc(Percentile)) |> slice(1)
low_metric <- scouting_df |> arrange(Percentile) |> slice(1)
```

<style>
body {
  background-color: #f7f9fc;
}

.report-hero {
  background: linear-gradient(135deg, #0b1f3a, #174a8b);
  color: white;
  padding: 32px;
  border-radius: 18px;
  margin-bottom: 24px;
}

.report-title {
  font-size: 38px;
  font-weight: 900;
  margin-bottom: 6px;
}

.report-subtitle {
  font-size: 18px;
  opacity: 0.95;
}

.kpi-grid {
  display: grid;
  grid-template-columns: repeat(5, 1fr);
  gap: 14px;
  margin-bottom: 28px;
}

.kpi-card {
  background: white;
  border-radius: 14px;
  padding: 18px;
  text-align: center;
  box-shadow: 0 4px 14px rgba(0,0,0,0.08);
}

.kpi-value {
  font-size: 28px;
  font-weight: 900;
  color: #0f172a;
}

.kpi-label {
  font-size: 12px;
  color: #64748b;
  font-weight: 800;
  text-transform: uppercase;
}

.section-card {
  background: white;
  border-radius: 16px;
  padding: 22px;
  margin-bottom: 24px;
  box-shadow: 0 4px 14px rgba(0,0,0,0.08);
}

.summary-box {
  background: #eef4ff;
  border-left: 6px solid #174a8b;
  padding: 18px;
  border-radius: 12px;
  font-size: 17px;
}
</style>

<div class="report-hero">
  <div class="report-title">`r player_name_param` Scouting Report</div>
  <div class="report-subtitle">`r player_type` profile | Team: `r player$team[1]` | Season: `r params$season`</div>
</div>

```{r}
if (player_type == "Hitter") {
  kpis <- tibble(
    Label = c("WAR", "wRC+", "OPS", "HR", "RBI"),
    Value = c(
      round(player$war, 1),
      round(player$w_rc_plus, 0),
      round(player$ops, 3),
      round(player$hr, 0),
      round(player$rbi, 0)
    )
  )
} else {
  kpis <- tibble(
    Label = c("WAR", "ERA", "FIP", "K/9", "BB/9"),
    Value = c(
      round(player$war, 1),
      round(player$era, 2),
      round(player$fip, 2),
      round(player$k_9, 2),
      round(player$bb_9, 2)
    )
  )
}

cat('<div class="kpi-grid">')
for (i in seq_len(nrow(kpis))) {
  cat(paste0(
    '<div class="kpi-card">',
    '<div class="kpi-value">', kpis$Value[i], '</div>',
    '<div class="kpi-label">', kpis$Label[i], '</div>',
    '</div>'
  ))
}
cat('</div>')
```

<div class="section-card">

## Player Overview

```{r}
overview_table <- player |>
  select(any_of(c(
    "player_name", "team", "pa", "ip", "hr", "rbi", "avg", "obp",
    "slg", "ops", "w_rc_plus", "era", "fip", "whip", "k_9", "bb_9", "war"
  )))

label_cols <- intersect(
  names(overview_table),
  c(
    "player_name", "team", "pa", "ip", "hr", "rbi", "avg", "obp",
    "slg", "ops", "w_rc_plus", "era", "fip", "whip", "k_9", "bb_9", "war"
  )
)

label_names <- c(
  player_name = "Player",
  team = "Team",
  pa = "PA",
  ip = "IP",
  hr = "HR",
  rbi = "RBI",
  avg = "AVG",
  obp = "OBP",
  slg = "SLG",
  ops = "OPS",
  w_rc_plus = "wRC+",
  era = "ERA",
  fip = "FIP",
  whip = "WHIP",
  k_9 = "K/9",
  bb_9 = "BB/9",
  war = "WAR"
)

overview_table |>
  gt() |>
  tab_header(
    title = md("**Player Stat Line**")
  ) |>
  cols_label(
    .list = as.list(label_names[label_cols])
  ) |>
  fmt_number(
    columns = where(is.numeric),
    decimals = 2
  ) |>
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    table.border.top.color = "#174A8B",
    table.border.bottom.color = "#174A8B"
  )
```

</div>

<div class="section-card">

## Percentile Profile

```{r}
ggplot(
  scouting_df,
  aes(x = reorder(Metric, Percentile), y = Percentile)
) +
  geom_col(fill = "#174A8B", width = 0.65) +
  geom_text(
    aes(label = paste0(Percentile, "%")),
    hjust = -0.15,
    fontface = "bold",
    size = 4
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(limits = c(0, 110), breaks = seq(0, 100, 25)) +
  labs(
    x = NULL,
    y = "Percentile"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_text(face = "bold"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(10, 40, 10, 10)
  )
```

```{r}
scouting_df |>
  gt() |>
  tab_header(
    title = md("**Scouting Metrics**")
  ) |>
  fmt_number(columns = c(Value, Percentile), decimals = 1) |>
  data_color(
    columns = Percentile,
    colors = col_numeric(
      palette = c("#fee2e2", "#fef3c7", "#dcfce7"),
      domain = c(0, 100)
    )
  ) |>
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    table.border.top.color = "#174A8B",
    table.border.bottom.color = "#174A8B"
  )
```

</div>

<div class="section-card">

## Summary

<div class="summary-box">
<strong>`r player_name_param`</strong> ranked highest in 
<strong>`r top_metric$Metric`</strong> with a percentile score of 
<strong>`r top_metric$Percentile`</strong>. The lowest relative area was 
<strong>`r low_metric$Metric`</strong> with a percentile score of 
<strong>`r low_metric$Percentile`</strong>.
</div>

</div>